---
title: "Intro to Statistical Programming"
output: ioslides_presentation
date: "Day 1, Data Science Basics in R Workshop"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Objects
An R object is the primary unit of "stuff" within R. 

An object could be:

  - a single number
  - a single word
  - a dataset
  - an entire policy document or set of documents
  - (among others)
  
## Create on object

To create an object in R, give it a name and assign it a value.

```{r object, echo = TRUE}
my_first_object <- 3.14
```

```{r objectprint, echo = TRUE}
my_first_object
```

## Data types

We will focus primarily on three related data types in this workshop: 

  - **numeric data:** numbers (e.g., -1.2, 0, 18)
  - **character:** strings of text (e.g., "malaria", "UNAIDS")
  - **logical:** TRUE or FALSE
  
## Numeric data

```{r numeric, echo = TRUE}
class(-1.2)
```

```{r numeric2, echo = TRUE}
class(c(-87.3, 8, 0))
```

## Character data

```{r character, echo = TRUE}
class("malaria")
```

```{r character2, echo = TRUE}
class(c("International Health Regulations", "UNAIDS", "34.2"))
```

## Logical data

```{r logical, echo = TRUE}
class(TRUE)
```

```{r logical2, echo = TRUE}
class(FALSE)
```
  
## Check your understanding

What type of data is this?
```{r test1, echo = TRUE, results = "hide"}
class(89.1)
```

## Check your understanding

What type of data is this?
```{r test1answer, echo = TRUE}
class(89.1)
```

## Data structures

We will focus primarily on two related data structures in this workshop: 

  - **vector:** a group of observations of a single data type
  - **data frames:** a data set made out of rows and columns that can contain different types of data 
  
## Vectors

```{r, echo = TRUE}
my_vector_numeric <- c(1, 43, -8, 9.26)
my_vector_numeric
```

```{r, echo = TRUE}
is(my_vector_numeric)
```

## Vectors

```{r, echo = TRUE}
my_vector_character <- c("Afghanistan", "Albania", "Algeria", "Andorra", "Angola", "Antigua and Barbuda")
my_vector_character
```

```{r, echo = TRUE}
is(my_vector_character)
```

## Data frames

  - Data frames are like "data sets"
  - Rows and columns of data
  - Often loaded from a file or from a website

## Data frames

```{r, echo = TRUE}
strep_tb <- read.csv("https://raw.githubusercontent.com/higgi13425/medicaldata/master/data-raw/strep_tb.csv")
```

```{r, echo = TRUE}
head(strep_tb[,1:6])
```
## Data frames contain vectors

```{r, echo = TRUE}
strep_tb$gender
```
```{r, echo = TRUE}
is(strep_tb$gender)
```

