---
title: "Intro to Statistical Programming"
output: ioslides_presentation
date: "Day 1, Data Science Basics in R Workshop"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R as a Calculator

```{r echo = TRUE}
1 + 1
```

```{r echo = TRUE}
8 - 10
```

```{r  echo = TRUE}
(32198 + 8943289)/12
```

## Use R as (an advanced) calculator

```{r  echo = TRUE}
log(1)
```

```{r  echo = TRUE}
sqrt(64)
```

```{r  echo = TRUE}
abs(-14)
```

## Save objects

```{r  echo = TRUE}
my_first_object <- 3.14

my_first_object
```


```{r echo = TRUE}
my_second_object <- 1+1

my_second_object
```

## Objects can be strings

```{r, echo = TRUE}
policy <- "International Health Regulations (IHR)"
```

```{r, echo = TRUE}
organization <- "UNAIDS"
```

## Data types

```{r, echo = TRUE}
is(-1)
```

```{r, echo = TRUE}
is(TRUE)
```

```{r, echo = TRUE}
is("What is this?")
```

## Use objects to do math

```{r, echo = TRUE}
my_first_object * my_second_object
```

```{r, echo = TRUE}
my_first_object/my_second_object
```

```{r, echo = TRUE}
my_first_object^my_second_object
```

## Vectors

```{r, echo = TRUE}
c("HIV", "malaria", "TB")
```

```{r, echo = TRUE}
c(1419, 4832, 10342)
```

```{r, echo = TRUE}
c("HIV", "malaria", 10342)
```

## Functions

```{r, echo = TRUE}
mean(c(1,2,3,4,5))
```

```{r, echo = TRUE}
sd(c(1,2,3,4,5))
```
```{r, echo = TRUE}
summary(c(1,2,3,4,5))
```

## Help

```{r, echo = TRUE}
?summary
```

## Vectorized calculations

```{r, echo = TRUE}
c(1,2,3,4,5) + 1
```

```{r, echo = TRUE}
c(1,2,3,4,5) * 2
```

```{r, echo = TRUE}
c(1,2,3,4,5) + c(8,0,0,0,0)
```

## Vector Errors

```{r, echo = TRUE}
c(1,2,3,4,5) + c(8,0)
```

## Data types
  - **numeric data:** numbers (e.g., -1.2, 0, 18)
  - **character:** strings of text (e.g., "malaria", "UNAIDS")
  - **logical:** TRUE or FALSE
  
```{r numeric, echo = TRUE}
class(-1.2)
```

```{r numeric2, echo = TRUE}
class(c(-87.3, 8, 0))
```

```{r character, echo = TRUE}
class("malaria")
```

```{r character2, echo = TRUE}
class(c("International Health Regulations", "UNAIDS", "34.2"))
```

```{r logical, echo = TRUE}
class(TRUE)
```

```{r logical2, echo = TRUE}
class(FALSE)
```
  
## Check your understanding

What type of data is this?
```{r test1, echo = TRUE, results = "hide"}
class(89.1)
```

## Check your understanding

What type of data is this?
```{r test1answer, echo = TRUE}
class(89.1)
```

## Data structures

  - **vector:** a group of observations of a single data type
  - **data frames:** a data set made out of rows and columns that can contain different types of data 
  
```{r, echo = TRUE}
my_vector_numeric <- c(1, 43, -8, 9.26)
my_vector_numeric
```

```{r, echo = TRUE}
is(my_vector_numeric)
```

```{r, echo = TRUE}
my_vector_character <- c("Afghanistan", "Albania", "Algeria", "Andorra", "Angola", "Antigua and Barbuda")
my_vector_character
```

```{r, echo = TRUE}
is(my_vector_character)
```


## Data frames

```{r, echo = TRUE}
strep_tb <- read.csv("https://raw.githubusercontent.com/higgi13425/medicaldata/master/data-raw/strep_tb.csv")
```

```{r, echo = TRUE}
head(strep_tb[,1:6])
```

```{r, echo = TRUE}
strep_tb$gender
```

```{r, echo = TRUE}
is(strep_tb$gender)
```

